---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  tidy.opts = list(width.cutoff = 60), tidy = TRUE,
  dpi = 150, fig.asp = 0.5, fig.width = 7, fig.retina = 1,
  out.width = "95%",
  warning = FALSE, message = FALSE
)
```

# miae

<!-- badges: start -->
<!-- badges: end -->

**miae** is an R package for multiple imputation through autoencoders built with **Torch**. It's currently under development.


## 1. Installation

You can install the current development version of miae from [GitHub](https://github.com/) with:
``` r
# install.packages("devtools")
#devtools::install_github("agnesdeng/miae")
```

``` {r}
library(miae)
```

## 2. Multiple imputation with denoising autoencoder with dropout
```{r}
#load the penguins dataset
library(palmerpenguins)
#5 variables have missing values
colSums(is.na(penguins))


midae.imputed<-midae(data=penguins,m = 5, epochs = 10, batch.size = 16,
                  split.ratio = 0.7, shuffle = TRUE,
                  input.dropout = 0.2, hidden.dropout = 0.5,
                  optimizer = "adamW", learning.rate = 0.0001, 
                  weight.decay = 0.002, momentum = 0,
                  encoder.structure = c(128, 64, 32), 
                  decoder.structure = c(32, 64, 128),
                  act = "elu", init.weight="xavier.uniform", 
                  scaler= "minmax",
                  verbose = TRUE, print.every.n = 1, 
                  save.model = FALSE, path = NULL)

#obtain the fifth imputed dataset
midae.imputed[[5]]


#show the imputed values for missing entries in the variable "bill_length_mm"
show_var(imputation.list=midae.imputed,var.name="bill_length_mm",original.data = penguins)

```

## 3. Multiple imputation with variational autoencoder

```{r}

mivae.imputed<-mivae(data=penguins,m = 5, epochs = 10, batch.size = 16,
                  split.ratio = 0.7, shuffle = TRUE,
                  input.dropout = 0.2, hidden.dropout = 0.5,
                  optimizer = "adamW", learning.rate = 0.0001, 
                  weight.decay = 0.002, momentum = 0,
                  encoder.structure = c(128, 64, 32), 
                  latent.dim = 4,
                  decoder.structure = c(32, 64, 128),
                  act = "elu", init.weight="xavier.uniform", 
                  scaler= "minmax",
                  verbose = FALSE, print.every.n = 1, 
                  save.model = FALSE, path = NULL)


#obtain the fifth imputed dataset
mivae.imputed[[5]]


#show the imputed values for missing entries in the variable "bill_length_mm"
show_var(imputation.list=mivae.imputed,var.name="bill_length_mm",original.data = penguins)
```
## 4. Impute new data using a saved imputation model

```{r}
n <- nrow(penguins)
idx <- sample(1:n, size = round(0.7 * n), replace = FALSE)
train.data <- penguins[idx, ]
test.data <- penguins[-idx, ]



midae.imputed<-midae(data=train.data,m = 5, epochs = 10,  
                  scaler = "minmax",
                  save.model=TRUE,
                  path=file.path(tempdir(),"midaemodel.pt"))


midae.newdata <- impute_new(path = file.path(tempdir(),"midaemodel.pt"),                               
                            newdata = test.data, scaler = "minmax",
                            m = 5)


```
