---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  tidy.opts = list(width.cutoff = 60), tidy = TRUE,
  dpi = 150, fig.asp = 0.5, fig.width = 7, fig.retina = 1,
  out.width = "95%",
  warning = FALSE, message = FALSE
)
```

# miae

<!-- badges: start -->
<!-- badges: end -->

**miae** is an R package for multiple imputation through autoencoders built with **Torch**. It's currently under development.


## 1. Installation

You can install the current development version of miae from [GitHub](https://github.com/) with:
``` r
# install.packages("devtools")
#devtools::install_github("agnesdeng/miae")
```

``` {r}
library(miae)
```

## 2. Multiple imputation with denoising autoencoder with dropout
```{r}
#load the penguins dataset
library(palmerpenguins)
#5 variables have missing values
colSums(is.na(penguins))


midae.imputed<-midae(data=penguins,m = 5, epochs = 10, batch.size = 50,
                  input.dropout = 0.9, latent.dropout = 0.5, hidden.dropout = 1,
                  optimizer = "adam", learning.rate = 0.001, weight.decay = 0, momentum = 0,
                  encoder.structure = c(128, 64, 32), latent.dim = 8, decoder.structure = c(32, 64, 128),
                  verbose = TRUE, print.every.n = 1, path = file.path(tempdir(),"midaemodel.pt"))

#obtain the fifth imputed dataset
midae.imputed[[5]]


#show the imputed values for missing entries in the variable "bill_length_mm"
show_var(imputation.list=midae.imputed,var.name="bill_length_mm",original.data = penguins)

```

## 3. Multiple imputation with variational autoencoder

```{r}
#load the penguins dataset
library(palmerpenguins)
#5 variables have missing values
colSums(is.na(penguins))


mivae.imputed<-mivae(data=penguins,m=5,epochs=10,path = file.path(tempdir(),"mivaemodel.pt"))

#obtain the fifth imputed dataset
mivae.imputed[[5]]


#show the imputed values for missing entries in the variable "bill_length_mm"
show_var(imputation.list=mivae.imputed,var.name="bill_length_mm",original.data = penguins)
```
## 4. Impute new data using a saved imputation model

```{r}
n <- nrow(penguins)
idx <- sample(1:n, size = round(0.7 * n), replace = FALSE)
train.data <- penguins[idx, ]
test.data <- penguins[-idx, ]



midae.data <- midae(data = train.data, m = 5, epochs = 5, path = "C:/Users/agnes/Desktop/torch/midaemodel.pt")
midae.newdata <- impute_new(path = "C:/Users/agnes/Desktop/torch/midaemodel.pt", newdata = test.data, m = 5)



mivae.data <- mivae(data = train.data, m = 5, epochs = 5, path = "C:/Users/agnes/Desktop/torch/mivaemodel.pt")
mivae.newdata <- impute_new(path = "C:/Users/agnes/Desktop/torch/mivaemodel.pt", newdata = test.data, m = 5)


```
