---
title: "Tuning dropout and visual diagnostics"
author: "Yongshi Deng"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tune-dropout}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy.opts = list(width.cutoff = 60), tidy = TRUE,
  dpi = 200, fig.asp = 0.5, fig.width = 7, fig.retina = 1,
  out.width = "95%",
  warning = FALSE, message = FALSE
)
```

# midae
```{r}
library(miae)
#devtools::load_all()
withNA.df <- createNA(data=iris,p=0.3)
midae.data <- midae(data=withNA.df, m = 5,
                  epochs = 5, batch.size = 16,
                  input.dropout = 0.8, latent.dropout = 0, hidden.dropout = 0.5,
                  optimizer = "adam", learning.rate = 0.001, weight.decay = 0, momentum = 0,
                  encoder.structure = c(256, 256, 256), latent.dim = 4, decoder.structure = c(256, 256, 256),
                  verbose = TRUE, print.every.n = 1)
plot_hist(imputation.list=midae.data,var.name="Sepal.Length",original.data=withNA.df)
plot_density(imputation.list=midae.data,var.name="Sepal.Length",original.data=withNA.df)

```
```{r}
library(mixgb)
mixgb.data <- mixgb(data=withNA.df, m = 5, maxit = 1)
plot_hist(imputation.list= mixgb.data,var.name="Sepal.Length",original.data=withNA.df)
plot_density(imputation.list=mixgb.data,var.name="Sepal.Length",original.data=withNA.df)
```
```{r}
library(mice)
mice.imp <- mice(data=withNA.df, m = 5, maxit = 5)
mice.data <- complete(mice.imp, action="all")
plot_hist(imputation.list= mice.data,var.name="Sepal.Length",original.data=withNA.df)
plot_density(imputation.list= mice.data,var.name="Sepal.Length",original.data=withNA.df)
```
It seems like the imputation variance of midae is not quite right. Let's tune the dropouts and show their differences.
```{r}
devtools::document()
devtools::load_all()
dropout.grid<-list(input.dropout=c(0.2,0.4,0.6,0.8),hidden.dropout=c(0.2, 0.4, 0.6, 0.8))

midae.tune<-tune_dae_dropout(data=withNA.df, m=5, epochs = 5, batch.size = 100,
                             dropout.grid = dropout.grid, latent.dropout = 0,
                             optim = "adam", learning.rate = 0.001, weight.decay = 0, momentum = 0,
                             encoder.structure = c(256, 256, 256), latent.dim = 4, decoder.structure = c(256, 256, 256),
                             verbose = TRUE, print.every.n = 1)
plot_dropout(tune.results = midae.tune,var.name="Sepal.Length")
```
Seems like some imputed values are too small or too large. Let's have a closer look. When `input_dropout` = 0.6 and `hidden_dropout` = 0.6 looks better. 
```{r}
range(iris$Sepal.Length)
plot_dropout(tune.results = midae.tune,var.name="Sepal.Length", xlim = range(iris$Sepal.Length) )
```

Now look at the other variables.

```{r}
plot_dropout(tune.results = midae.tune,var.name="Sepal.Width", xlim = range(iris$Sepal.Width) )
plot_dropout(tune.results = midae.tune,var.name="Petal.Length", xlim = range(iris$Petal.Length) )
plot_dropout(tune.results = midae.tune,var.name="Petal.Width", xlim = range(iris$Petal.Width) )
```
Use the tuned dropout to impute.
```{r}
midae.data1 <- midae(data=withNA.df, m = 5,
                  epochs = 5, batch.size = 16,
                  input.dropout = 0.6, latent.dropout = 0, hidden.dropout = 0.6,
                  optimizer = "adam", learning.rate = 0.001, weight.decay = 0, momentum = 0,
                  encoder.structure = c(256, 256, 256), latent.dim = 4, decoder.structure = c(256, 256, 256),
                  verbose = TRUE, print.every.n = 1)
plot_hist(imputation.list=midae.data1,var.name="Sepal.Length",original.data=withNA.df)
plot_density(imputation.list=midae.data1,var.name="Sepal.Length",original.data=withNA.df)
```
```{r}
midae.data2 <- midae(data=withNA.df, m = 5,
                  epochs = 5, batch.size = 16,
                  input.dropout = 0.8, latent.dropout = 0, hidden.dropout = 0.2,
                  optimizer = "adam", learning.rate = 0.001, weight.decay = 0, momentum = 0,
                  encoder.structure = c(256, 256, 256), latent.dim = 4, decoder.structure = c(256, 256, 256),
                  verbose = TRUE, print.every.n = 1)
plot_hist(imputation.list=midae.data2,var.name="Sepal.Length",original.data=withNA.df)
plot_density(imputation.list=midae.data2,var.name="Sepal.Length",original.data=withNA.df)
```
```{r}
#midas's default  input.dropout=0.8, hidden.dropout=0.5
plot_density(imputation.list=midae.data,var.name="Sepal.Length",original.data=withNA.df)
#input.dropout=0.6, hidden.dropout=0.6
plot_density(imputation.list=midae.data1,var.name="Sepal.Length",original.data=withNA.df)
#input.dropout=0.8, hidden.dropout=0.2
plot_density(imputation.list=midae.data2,var.name="Sepal.Length",original.data=withNA.df)
```


# mivae
Under the default settings of `rmidas`.
```{r}
mivae.data <- mivae(data=withNA.df, m = 5,
                  epochs = 5, batch.size = 16,
                  input.dropout = 0.8, latent.dropout = 0, hidden.dropout = 0.5,
                  optimizer = "adam", learning.rate = 0.001, weight.decay = 0, momentum = 0,
                  encoder.structure = c(256, 256, 256), latent.dim = 4, decoder.structure = c(256, 256, 256),
                  verbose = TRUE, print.every.n = 1)
plot_hist(imputation.list=mivae.data,var.name="Sepal.Length",original.data=withNA.df)
plot_density(imputation.list=mivae.data,var.name="Sepal.Length",original.data=withNA.df)
```




```{r}
dropout.grid<-list(input.dropout=c(0.2,0.4,0.6,0.8),hidden.dropout=c(0.2, 0.4, 0.6, 0.8))

mivae.tune<-tune_vae_dropout(data=withNA.df, m=5, epochs = 5, batch.size = 16,
                             dropout.grid = dropout.grid, latent.dropout = 0,
                             optim = "adam", learning.rate = 0.001, weight.decay = 0, momentum = 0,
                             encoder.structure = c(256, 256, 256), latent.dim = 4, decoder.structure = c(256, 256, 256),
                             verbose = TRUE, print.every.n = 1)
plot_dropout(tune.results = mivae.tune,var.name="Sepal.Length")
plot_dropout(tune.results = mivae.tune,var.name="Sepal.Length",xlim = range(iris$Sepal.Length))
```
Use the tune dropout to impute
```{r}
mivae.data1 <- mivae(data=withNA.df, m = 5,
                  epochs = 5, batch.size = 16,
                  input.dropout = 0.2, latent.dropout = 0, hidden.dropout = 0.8,
                  optimizer = "adam", learning.rate = 0.001, weight.decay = 0, momentum = 0,
                  encoder.structure = c(256, 256, 256), latent.dim = 4, decoder.structure = c(256, 256, 256),
                  verbose = TRUE, print.every.n = 1)
plot_hist(imputation.list=mivae.data1,var.name="Sepal.Length",original.data=withNA.df)
plot_density(imputation.list=mivae.data1,var.name="Sepal.Length",original.data=withNA.df)
plot_density(imputation.list=mivae.data,var.name="Sepal.Length",original.data=withNA.df)
```

```{r}
plot_density(imputation.list=midae.data1,var.name="Sepal.Length",original.data=withNA.df, xlim=range(iris$Sepal.Length))
plot_density(imputation.list=mivae.data1,var.name="Sepal.Length",original.data=withNA.df, xlim=range(iris$Sepal.Length))
plot_density(imputation.list=mixgb.data,var.name="Sepal.Length",original.data=withNA.df, xlim=range(iris$Sepal.Length))
plot_density(imputation.list=mice.data,var.name="Sepal.Length",original.data=withNA.df, xlim=range(iris$Sepal.Length))
```
