---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  tidy.opts = list(width.cutoff = 60), tidy = TRUE,
  dpi = 150, fig.asp = 0.5, fig.width = 7, fig.retina = 1,
  out.width = "95%",
  warning = FALSE, message = FALSE
)
```

# miae

<!-- badges: start -->
<!-- badges: end -->

**miae** is an R package for multiple imputation through autoencoders built with **Torch**. It's currently under development.


## 1. Installation

You can install the current development version of miae from [GitHub](https://github.com/) with:
``` r
# install.packages("devtools")
#devtools::install_github("agnesdeng/miae")
```

``` {r}
library(miae)
```

## 2. Multiple imputation with denoising autoencoder with dropout
```{r}
# load the penguins dataset
library(palmerpenguins)
# 5 variables have missing values
colSums(is.na(penguins))
```

```{r}
midae.imputed <- midae(
  data = penguins, m = 5, device = "cpu",
  pmm.type = "auto", pmm.k = 5, pmm.link = "prob", pmm.save.vars = NULL,
  epochs = 10, batch.size = 32,
  subsample = 0.7, shuffle = TRUE,
  input.dropout = 0.2, hidden.dropout = 0.5,
  optimizer = "adamW", learning.rate = 0.0001, weight.decay = 0.002, momentum = 0, eps = 1e-07,
  encoder.structure = c(128, 64, 32), decoder.structure = c(32, 64, 128),
  act = "elu", init.weight = "xavier.normal", scaler = "minmax",
  loss.na.scale = FALSE,
  early_stopping_epochs = 10,
  verbose = TRUE, print.every.n = 1, save.model = FALSE, path = file.path(tempdir(), "midaemodel.pt")
)

# obtain the fifth imputed dataset
midae.imputed[[5]]


# show the imputed values for missing entries in the variable "bill_length_mm"
show_var(imputation.list = midae.imputed, var.name = "bill_length_mm", original.data = penguins)
```

## 3. Multiple imputation with variational autoencoder

```{r}
mivae.imputed <- mivae(
  data = penguins, m = 5,
  beta = 1, pmm.type = "auto", pmm.k = 5, pmm.link = "prob", pmm.save.vars = NULL,
  epochs = 10, batch.size = 32,
  subsample = 0.7, shuffle = TRUE,
  input.dropout = 0.2, hidden.dropout = 0.5,
  optimizer = "adamW", learning.rate = 0.0001, weight.decay = 0.002, momentum = 0, eps = 1e-07,
  encoder.structure = c(128, 64, 32), latent.dim = 4, decoder.structure = c(32, 64, 128),
  act = "elu", init.weight = "xavier.normal", scaler = "minmax",
  loss.na.scale = FALSE,
  early_stopping_epochs = 10,
  verbose = TRUE, print.every.n = 1, save.model = FALSE, path = file.path(tempdir(), "mivaemodel.pt")
)


# obtain the fifth imputed dataset
mivae.imputed[[5]]


# show the imputed values for missing entries in the variable "bill_length_mm"
show_var(imputation.list = mivae.imputed, var.name = "bill_length_mm", original.data = penguins)
```

## 4. Impute new data using a saved imputation model

```{r}
n <- nrow(penguins)
idx <- sample(1:n, size = round(0.7 * n), replace = FALSE)
train.data <- penguins[idx, ]
test.data <- penguins[-idx, ]


midae.obj <- midae(
  data = train.data, m = 5, epochs = 10,
  subsample = 1,
  early_stopping_epochs = 1,
  scaler = "minmax",
  save.model = TRUE,
  path = file.path(tempdir(), "midaemodel.pt")
)


midae.newdata <- impute_new(object = midae.obj, newdata = test.data, m = 5)
```
